\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-04-24 doc/paper.tex
added refs for experimental results and praragraph selling FS.

  For instance from the recent $7$ TeV and $8$ TeV runs
at the Large Hadron Collider (LHC) there have been two important
developments.  First low energy signatures expected from such models,
such as the classic jets plus missing transverse energy signature,
have not been observed, substantially raising the lower limit on
sparticle masses\cite{}. No other signature of beyond the standard
model(BSM) physics has been observed, leaving the fundamental
questions which motivated BSM physics unanswered.

 Secondly ATLAS and CMS discovered\cite{} a light Higgs of $125$ GeV,
 within the mass range that could be accommodated in the MSSM but
 requiring stops which are significantly heavier than both the direct
 collider limits and indirect limits that appears in constrained
 models from the significantly higher limits on first and second
 generation squarks.  Alternatively it is possible that a $125$ GeV
 Higgs in non-minimal models may not require such heavy stops, indeed
 a number of such models have been constructed to do exactly this by
 having new tree level contributions to the Higgs mass.

These results lead to a number of different perspectives, for example
some may wish to abandon naturalness as a motivation and focus on BSM
models which can fit the observed relic density and allow unification
of gauge couplings, while others may focus on special models which
preserve naturalness.  In both cases exploration of such models can be
aided if it is possible to quickly create a fast spectrum generator.

\subsection{The Program}
\fs is a MATHEMATICA package designed to create a fast and easily
adaptable spectrum generator in C++ for any SUSY model and some
non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
existing \sarah package \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} to obtain one-loop self energies, one-loop
tadpoles corrections and two-loop renormalisation group
equations(RGEs) in any model.  \fs then converts these algerbraic
expressions from \sarah into C++ code using the C++11 standard. \fs
uses some parts of SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen
library \cite{eigen}, as well as the gnu scientific library and the
boost library to create numerical routines which then use these
routines to solve the RGEs and EWSB conditions which will be used in
the spectrum generator. The main solver routine of \fs then calls
these routines to find simultaneous solutions for the EWSB, low energy
data, and other, user supplied, boundary conditions (e.g. the usual
mSUGRA universality conditions) and calculates the mass spectrum.

@@

%\cite{Baer:1993ae}
\bibitem{Baer:1993ae} 
  H.~Baer, F.~E.~Paige, S.~D.~Protopopescu and X.~Tata,
  %``Simulating Supersymmetry with ISAJET 7.0 / ISASUSY 1.0,''
  hep-ph/9305342.
  %%CITATION = HEP-PH/9305342;%%
  %82 citations counted in INSPIRE as of 03 Nov 2013
%\cite{Ellwanger:2006rn}
\bibitem{Ellwanger:2006rn} 
  U.~Ellwanger and C.~Hugonie,
  %``NMSPEC: A Fortran code for the sparticle and Higgs masses in the NMSSM with GUT scale boundary conditions,''
  Comput.\ Phys.\ Commun.\  {\bf 177}, 399 (2007)
  [hep-ph/0612134].
  %%CITATION = HEP-PH/0612134;%%
  %81 citations counted in INSPIRE as of 03 Nov 2013
%\cite{Allanach:2013kza}
\bibitem{Allanach:2013kza} 
  B.~C.~Allanach, P.~Athron, L.~C.~Tunstall, A.~Voigt and A.~G.~Williams,
  %``Next-to-Minimal SOFTSUSY,''
  arXiv:1311.7659 [hep-ph].
  %%CITATION = ARXIV:1311.7659;%%
  %1 citations counted in INSPIRE as of 24 Apr 2014
  %\cite{Staub:2010ty}
\bibitem{NMSSM} P. Fayet, Nucl. Phys. B \textbf{90} (1975) 104; Phys. Lett.
B \textbf{64} (1976) 159; Phys. Lett. B \textbf{69} (1977) 489 and Phys. Lett. B
\textbf{84} (1979) 416; H.P. Nilles, M. Srednicki and D. Wyler, Phys. Lett. B
\textbf{120} (1983) 346; J.M. Frere, D.R. Jones and S. Raby, Nucl. Phys. B
\textbf{222} (1983) 11; J.P. Derendinger and C.A. Savoy, Nucl. Phys. B
\textbf{237} (1984) 307;  A.I. Veselov, M.I. Vysotsky and K.A. Ter-Martirosian,
Sov. Phys. JETP \textbf{63} (1986) 489; J.R. Ellis, J.F. Gunion, H.E. Haber, L.
Roszkowski and F. Zwirner, Phys. Rev. D \textbf{39}  (1989) 844; M. Drees, Int.
J. Mod. Phys. A \textbf{4}  (1989) 3635; U. Ellwanger, M. Rausch de
Traubenberg and C.A. Savoy, Phys. 
Lett. B \textbf{315} (1993) 331, Z. Phys. C {\bf 67} (1995) 665 and Nucl. Phys.
B \textbf{492} (1997) 307; U.~Ellwanger, Phys.\ Lett.\  B {\bf 303} (1993) 271; P.
Pandita, Z. Phys. C \textbf{59} (1993) 575; T. Elliott, S.F. King and P.L.
White, Phys. Rev. D {\bf 49} (1994) 2435; S.F. King and P.L. White, Phys. Rev. D
\textbf{52} (1995) 4183;  F.~Franke and H.~Fraas, Int.\ J.\ Mod.\ Phys.\  A {\bf
12} (1997) 479.   D.~J.~Miller, R.~Nevzorov and P.~M.~Zerwas,  Nucl.\ Phys.\ B {\bf 681}, 3 (2004) [hep-ph/0304049].
%\cite{Kim:1983dt}
\bibitem{Kim:1983dt} 
  J.~E.~Kim and H.~P.~Nilles,
  %``The mu Problem and the Strong CP Problem,''
  Phys.\ Lett.\ B {\bf 138}, 150 (1984).
  %%CITATION = PHLTA,B138,150;%%
  %604 citations counted in INSPIRE as of 24 Apr 2014
%\cite{King:2014nza}
\bibitem{King:2014nza} 
  S.~F.~King, A.~Merle, S.~Morisi, Y.~Shimizu and M.~Tanimoto,
  %``Neutrino Mass and Mixing: from Theory to Experiment,''
  arXiv:1402.4271 [hep-ph].
  %%CITATION = ARXIV:1402.4271;%%
  %9 citations counted in INSPIRE as of 24 Apr 2014
%\cite{King:2008qb}
\bibitem{King:2008qb} 
  S.~F.~King, R.~Luo, D.~J.~Miller and R.~Nevzorov,
  %``Leptogenesis in the Exceptional Supersymmetric Standard Model: Flavour dependent lepton asymmetries,''
  JHEP {\bf 0812}, 042 (2008)
  [arXiv:0806.0330 [hep-ph]].
  %%CITATION = ARXIV:0806.0330;%%
  %26 citations counted in INSPIRE as of 24 Apr 2014
\bibitem{Staub:2010ty} 
  F.~Staub, W.~Porod and B.~Herrmann,
  %``The Electroweak sector of the NMSSM at the one-loop level,''
  JHEP {\bf 1010}, 040 (2010)
  [arXiv:1007.4049 [hep-ph]].
  %%CITATION = ARXIV:1007.4049;%%
  %28 citations counted in INSPIRE as of 12 Oct 2013

>>>>

Recent experimental developments have also increased the relevancy of
such a tool. From the recent $7$ TeV and $8$ TeV runs at the Large
Hadron Collider (LHC) there have been two important developments.
First low energy signatures expected from such models, such as the
classic jets plus missing transverse energy signature, have not been
observed, substantially raising the lower limit on sparticle masses
(see e.g.~\cite{Aad:2013wta,Chatrchyan:2014lfa}). No other signature
of beyond the standard model(BSM) physics has been observed, leaving
the fundamental questions which motivated BSM physics unanswered.

 Secondly ATLAS and CMS discovered\cite{ATLAS:2012ae,
   Chatrchyan:2012tx} a light Higgs of $125$ GeV, within the mass
 range that could be accommodated in the MSSM but requiring stops
 which are significantly heavier than both the direct collider limits
 and indirect limits that appears in constrained models from the
 significantly higher limits on first and second generation squarks.
 Alternatively it is possible that a $125$ GeV Higgs in non-minimal
 models may not require such heavy stops, indeed a number of such
 models have been constructed to do exactly this by having new tree
 level contributions to the Higgs mass.

These motivate both exploring non-minimal SUSY models which ameliorate
the naturaless problems introduced, by, for example, raising the tree
level Higgs mass, and models developed with a fresh persp[ective,
  based on other considerations.  In both cases exploration of such
  models can be aided if it is possible to quickly create a fast
  spectrum generator.

\fs is designed for precisely this task. \fs is a MATHEMATICA package
which with the help of \sarah
\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
creates a fast, modular C++ spectrum generator for a user specified
SUSY model.  The generated code structure is designed to be as
flexible as possible to accomodate different types of extensions and
due to it's modular nature it is easy to modify, add new featuires to
or combine with other programs.

\subsection{The Program}
\label{Sec:Program}
\fs is a MATHEMATICA package designed to create a fast and easily
adaptable spectrum generator in C++ for any SUSY model and some
non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
existing \sarah package \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} to obtain one-loop self energies, one-loop
tadpoles corrections and two-loop renormalisation group
equations(RGEs) in any model.  \fs then converts these algerbraic
expressions from \sarah into C++ code using the C++11 standard. \fs
uses some parts of SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen
library \cite{eigen}, as well as the gnu scientific library and the
boost library to create numerical routines which then use these
routines to solve the RGEs and EWSB conditions which will be used in
the spectrum generator. The main solver routine of \fs then calls
these routines to find simultaneous solutions for the EWSB, low energy
data, and other, user supplied, boundary conditions (e.g. the usual
mSUGRA universality conditions) and calculates the mass spectrum.

@@

%\cite{Baer:1993ae}
\bibitem{Baer:1993ae} 
  H.~Baer, F.~E.~Paige, S.~D.~Protopopescu and X.~Tata,
  %``Simulating Supersymmetry with ISAJET 7.0 / ISASUSY 1.0,''
  hep-ph/9305342.
  %%CITATION = HEP-PH/9305342;%%
  %82 citations counted in INSPIRE as of 03 Nov 2013
%\cite{Ellwanger:2006rn}
\bibitem{Ellwanger:2006rn} 
  U.~Ellwanger and C.~Hugonie,
  %``NMSPEC: A Fortran code for the sparticle and Higgs masses in the NMSSM with GUT scale boundary conditions,''
  Comput.\ Phys.\ Commun.\  {\bf 177}, 399 (2007)
  [hep-ph/0612134].
  %%CITATION = HEP-PH/0612134;%%
  %81 citations counted in INSPIRE as of 03 Nov 2013
%\cite{Allanach:2013kza}
\bibitem{Allanach:2013kza} 
  B.~C.~Allanach, P.~Athron, L.~C.~Tunstall, A.~Voigt and A.~G.~Williams,
  %``Next-to-Minimal SOFTSUSY,''
  arXiv:1311.7659 [hep-ph].
  %%CITATION = ARXIV:1311.7659;%%
  %1 citations counted in INSPIRE as of 24 Apr 2014
  %\cite{Staub:2010ty}
\bibitem{NMSSM} P. Fayet, Nucl. Phys. B \textbf{90} (1975) 104; Phys. Lett.
B \textbf{64} (1976) 159; Phys. Lett. B \textbf{69} (1977) 489 and Phys. Lett. B
\textbf{84} (1979) 416; H.P. Nilles, M. Srednicki and D. Wyler, Phys. Lett. B
\textbf{120} (1983) 346; J.M. Frere, D.R. Jones and S. Raby, Nucl. Phys. B
\textbf{222} (1983) 11; J.P. Derendinger and C.A. Savoy, Nucl. Phys. B
\textbf{237} (1984) 307;  A.I. Veselov, M.I. Vysotsky and K.A. Ter-Martirosian,
Sov. Phys. JETP \textbf{63} (1986) 489; J.R. Ellis, J.F. Gunion, H.E. Haber, L.
Roszkowski and F. Zwirner, Phys. Rev. D \textbf{39}  (1989) 844; M. Drees, Int.
J. Mod. Phys. A \textbf{4}  (1989) 3635; U. Ellwanger, M. Rausch de
Traubenberg and C.A. Savoy, Phys. 
Lett. B \textbf{315} (1993) 331, Z. Phys. C {\bf 67} (1995) 665 and Nucl. Phys.
B \textbf{492} (1997) 307; U.~Ellwanger, Phys.\ Lett.\  B {\bf 303} (1993) 271; P.
Pandita, Z. Phys. C \textbf{59} (1993) 575; T. Elliott, S.F. King and P.L.
White, Phys. Rev. D {\bf 49} (1994) 2435; S.F. King and P.L. White, Phys. Rev. D
\textbf{52} (1995) 4183;  F.~Franke and H.~Fraas, Int.\ J.\ Mod.\ Phys.\  A {\bf
12} (1997) 479.   D.~J.~Miller, R.~Nevzorov and P.~M.~Zerwas,  Nucl.\ Phys.\ B {\bf 681}, 3 (2004) [hep-ph/0304049].
%\cite{Kim:1983dt}
\bibitem{Kim:1983dt} 
  J.~E.~Kim and H.~P.~Nilles,
  %``The mu Problem and the Strong CP Problem,''
  Phys.\ Lett.\ B {\bf 138}, 150 (1984).
  %%CITATION = PHLTA,B138,150;%%
  %604 citations counted in INSPIRE as of 24 Apr 2014
%\cite{King:2014nza}
\bibitem{King:2014nza} 
  S.~F.~King, A.~Merle, S.~Morisi, Y.~Shimizu and M.~Tanimoto,
  %``Neutrino Mass and Mixing: from Theory to Experiment,''
  arXiv:1402.4271 [hep-ph].
  %%CITATION = ARXIV:1402.4271;%%
  %9 citations counted in INSPIRE as of 24 Apr 2014
%\cite{King:2008qb}
\bibitem{King:2008qb} 
  S.~F.~King, R.~Luo, D.~J.~Miller and R.~Nevzorov,
  %``Leptogenesis in the Exceptional Supersymmetric Standard Model: Flavour dependent lepton asymmetries,''
  JHEP {\bf 0812}, 042 (2008)
  [arXiv:0806.0330 [hep-ph]].
  %%CITATION = ARXIV:0806.0330;%%
  %26 citations counted in INSPIRE as of 24 Apr 2014
%\cite{Aad:2013wta}
\bibitem{Aad:2013wta} 
  G.~Aad {\it et al.}  [ATLAS Collaboration],
  %``Search for new phenomena in final states with large jet multiplicities and missing transverse momentum at sqrt(s)=8 TeV proton-proton collisions using the ATLAS experiment,''
  JHEP {\bf 1310}, 130 (2013)
  [arXiv:1308.1841 [hep-ex]].
  %%CITATION = ARXIV:1308.1841;%%
  %36 citations counted in INSPIRE as of 24 Apr 2014
\bibitem{Chatrchyan:2014lfa}
  S.~Chatrchyan {\it et al.}  [CMS Collaboration],
  %``Search for new physics in the multijet and missing transverse momentum final state in proton-proton collisions at $\sqrt{s}$ = 8 TeV,''
  arXiv:1402.4770 [hep-ex].
  %%CITATION = ARXIV:1402.4770;%%
  %10 citations counted in INSPIRE as of 24 Apr 2014
%\cite{ATLAS:2012ae}
\bibitem{ATLAS:2012ae} 
  G.~Aad {\it et al.}  [ATLAS Collaboration],
  %``Combined search for the Standard Model Higgs boson using up to 4.9 fb$^{-1}$ of $pp$ collision data at $\sqrt{s}=7$ TeV with the ATLAS detector at the LHC,''
  Phys.\ Lett.\ B {\bf 710}, 49 (2012)
  [arXiv:1202.1408 [hep-ex]].
  %%CITATION = ARXIV:1202.1408;%%
  %474 citations counted in INSPIRE as of 24 Apr 2014
%\cite{Chatrchyan:2012tx}
\bibitem{Chatrchyan:2012tx} 
  S.~Chatrchyan {\it et al.}  [CMS Collaboration],
  %``Combined results of searches for the standard model Higgs boson in $pp$ collisions at $\sqrt{s}=7$ TeV,''
  Phys.\ Lett.\ B {\bf 710}, 26 (2012)
  [arXiv:1202.1488 [hep-ex]].
  %%CITATION = ARXIV:1202.1488;%%
  %592 citations counted in INSPIRE as of 24 Apr 2014
%\cite{Chatrchyan:2014lfa}
\bibitem{Staub:2010ty} 
  F.~Staub, W.~Porod and B.~Herrmann,
  %``The Electroweak sector of the NMSSM at the one-loop level,''
  JHEP {\bf 1010}, 040 (2010)
  [arXiv:1007.4049 [hep-ph]].
  %%CITATION = ARXIV:1007.4049;%%
  %28 citations counted in INSPIRE as of 12 Oct 2013
